---
title:  "Shiny/Leaflet"
subtitle: "Building a USA water dashboard"
output:
  html_document:
    toc: true
    toc_float: true
    toc_collapsed: true
---

```{r,echo=FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE, comment = "#>", out.width = "100%")
```

# Libraries

```{r}
library(AOI)           # Data and geocoding
library(dataRetrieval) # USGS data access

library(sf)            # all things spatial ...
library(dplyr)         # data manipulations ...

library(leaflet)       # mapping
```

# Data

### County `sf` object 

```{r}
counties = aoi_get(state = 'conus', county = "all") %>%
  mutate(location = paste(name, state_abbr, sep = ", ")) %>% 
  select(geoid, name, state_name, location)

glimpse(counties)
plot(counties$geometry)
```

### Active USGS gages

```{r}
sites   = readRDS("shiny/usgs_sites.rds")
glimpse(sites)
```

# Core Leaflet functions

### Auto generate basemap

```{r}
basemap = function(){
  leaflet() %>%
    addProviderTiles('CartoDB.Positron') %>%
    setView(lat = 39, lng = -95, zoom = 3)
}
```

#### Example 

```{r}
basemap()
```

### Zoom to a county, given a GEOID

```{r}
zoom_to_county = function(map, counties, FIP){

  shp    = filter(counties, geoid == FIP)
  bounds = as.vector(st_bbox(shp))

  clearGroup(map, 'shp') %>%
    addPolygons(data = shp,
                color      = "#003660",
                fillColor  = "#FEBC11",
                fillOpacity = .2,
                group = "shp") %>%
    flyToBounds(bounds[1], bounds[2], bounds[3], bounds[4])
}
```

#### Example 

```{r}
zoom_to_county(basemap(), counties, 25017)
```

# USGS functions

### Wrapper for finding streamflow given a siteID

```{r}
get_streamflow = function(siteID){
  readNWISdv(siteNumbers = siteID,
             parameterCd = '00060',
             startDate   = Sys.Date() - 365) %>%
    renameNWISColumns()
}
```

#### Example 

```{r}
flow = get_streamflow('01098530')
glimpse(flow)
```

### [Network Linked Data Index](https://usgs-r.github.io/dataRetrieval/articles/nldi.html) function given a lat/lon

```{r}
find_nldi  = function(x, y){
  findNLDI(location = c(x,y), nav = c('UM'), 
           find = c("basin", "flowlines"), distance = 1000)
}
```

#### Example 

```{r}
nldi_out = find_nldi(-104.780837, 38.786796)

basemap() %>% 
  addPolygons(data = nldi_out$basin) %>% 
  addPolygons(data = nldi_out$UM_flowlines) %>%
  zoom_to_county(counties, '08041')
```

# Where are we going?

We will use [flexdashboard](https://rmarkdown.rstudio.com/flexdashboard/) to generate the app. 

If you know Rmarkdown basics, this will be intuitive for you!

The key new piece is that:

 - Each Level 1 Header (#) begins a new page in the dashboard.
 - Each Level 2 Header (##) begins a new column/row.
 - Each Level 3 Header (###) begins a new box.

The app material can be found [here](https://github.com/mikejohnson51/leaflet-intro/tree/master/shiny)
